










                            ┌─────────────────────────────────────────────────────────────┐
                            │                    JavaScript Runtime                       │
                            │                                                             │
                            │  ┌─────────────────┐  ┌─────────────────┐  ┌─────────────┐  │
                            │  │   CALL STACK    │  │   WEB APIs      │  │   QUEUE     │  │
                            │  │  (Synchronous)  │  │  (Asynchronous) │  │ (Callbacks) │  │
                            │  │                 │  │                 │  │             │  │
                            │  │ • console.log   │  │ • setTimeout    │  │ • Callbacks │  │
                            │  │ • calculations  │  │ • fetch/XHR     │  │ • Promises  │  │
                            │  │ • function calls│  │ • DOM events    │  │ • Microtasks│  │
                            │  │                 │  │ • I/O operations│  │             │  │
                            │  └─────────────────┘  └─────────────────┘  └─────────────┘  │
                            │          │                         │              │         │
                            │          └─────────────────────────┼──────────────┘         │
                            │                                    │                        │
                            │                            EVENT LOOP ◄─────────────────────┘
                            │              (Checks: Call Stack empty? → Move from Queue)
                            │
                            └─────────────────────────────────────────────────────────────┘







// ============================================================
// JAVASCRIPT RUNTIME - MAIN POINTS ONLY
// ============================================================


// MAIN POINT 1: CALL STACK (Synchronous - waits for each step)
// Code runs line by line, one at a time


console.log("1. First");


function greet(name) {
console.log("Hello, " + name);
}


greet("Ali");        // Waits for this to finish
console.log("2. Last");


// Output:
// 1. First
// Hello, Ali
// 2. Last


// ============================================================
// MAIN POINT 2: WEB APIs (Asynchronous - does not wait)
// Browser handles slow tasks in background, code continues


console.log("Start");


setTimeout(() => {
console.log("After 2 seconds");
}, 2000);


console.log("End");  // Runs immediately, does not wait!


// Output:
// Start
// End
// After 2 seconds (arrives later)


// ============================================================
// MAIN POINT 3: QUEUE (Where callbacks wait)
// When Web APIs finish, callbacks wait in queue


console.log("A");


setTimeout(() => {
console.log("B - Waiting in queue");
}, 0);  // Even 0 milliseconds goes to queue!


console.log("C");


// Output:
// A
// C
// B - Waiting in queue


// ============================================================
// MAIN POINT 4: EVENT LOOP (Checks: Is Call Stack empty?)
// If empty, moves callbacks from queue to call stack


// Example showing Event Loop in action:


console.log("------- EVENT LOOP EXAMPLE -------");


console.log("1. Start");


// Web API - goes to queue after 1 second
setTimeout(() => {
console.log("3. From queue to stack");
}, 1000);


console.log("2. End");


// Event Loop steps:
// Step 1: Call Stack runs "1. Start"
// Step 2: Call Stack runs "2. End"
// Step 3: Call Stack is empty now
// Step 4: Event Loop checks - "Is Call Stack empty?"
// Step 5: YES! Event Loop takes callback from queue
// Step 6: Puts callback in Call Stack
// Step 7: Call Stack runs "3. From queue to stack"


// Output:
// ------- EVENT LOOP EXAMPLE -------
// 1. Start
// 2. End
// 3. From queue to stack


// ============================================================
// COMPLETE EXAMPLE - ALL PARTS WORKING TOGETHER
// ============================================================


console.log("===== COMPLETE EXAMPLE =====");


// Call Stack - runs immediately
console.log("A");


// Web API - setTimeout (goes to queue after delay)
setTimeout(() => {
console.log("B");
}, 1000);


// Call Stack - runs immediately
console.log("C");


// Web API - Promise (goes to queue, high priority)
Promise.resolve().then(() => {
console.log("D");
});


// Call Stack - runs immediately
console.log("E");


// EXECUTION:
// Call Stack runs: A → C → E
// Event Loop checks: Call Stack empty?
// YES! Move queue items to stack
// Promises run first (high priority): D
// Then setTimeout: B


// Output:
// ===== COMPLETE EXAMPLE =====
// A
// C
// E
// D
// B


// ============================================================
// KEY CONCEPT: Why JavaScript does not freeze?
// ============================================================


// WITHOUT Event Loop (would freeze):
// console.log("Start");
// wait for 10 seconds... // Program completely frozen!
// console.log("Done");   // Waits 10 seconds


// WITH Event Loop (does not freeze):
console.log("Start fetching");


// This does not freeze the program!
fetch("https://jsonplaceholder.typicode.com/posts/1")
.then(response => response.json())
.then(data => console.log("Got data:", data.title));


console.log("Program continues, not frozen!");


// Output:
// Start fetching
// Program continues, not frozen!
// Got data: sunt aut facere repellat (arrives later)





































